### 2015-09-09 fundamentals/impl.tex
fix sim section typo
perhaps this is the right image?
remove apparently unrelated comments
Reworking of QA section

\begin{figure}[htbp!]
\begin{center}
\includegraphics{./images/modifiedopen.eps}
\end{center}
\caption{The \Cyclus framework enables fully open, partially open, and fully
closed collaborations\cite{carlsen_cyclus_2014}.}
\label{fig:modifiedopen}
\end{figure}

@@

\begin{figure}[htbp!]
\begin{center}
\includegraphics{./images/compositions}
\end{center}
\caption{
A simple decay line cache holding compositions A, B, C, and a
yet-uncomputed composition D.  B comes from decaying A 1 time step.  C comes
from decaying B 2 time steps, etc.  Black represents the cache for this
particular composition chain. Blue indicates decay operations that can be
satisfied by cache lookups.   If A needs to be decayed 1 time step (A +1), a
quick lookup returns the previously computed B.  Decaying B 3 time steps
requires a decay calculation to compute a new composition D, but subsequent
requests such as decaying A 4 time steps will not require any recalculation.
}
\label{fig:compositions}
\end{figure}

@@

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|r|l|X|}
\hline
\textbf{Entity} & \textbf{Archetype} & \textbf{Functionality} \\
\hline
Facility & EnrichmentFacility & This facility enriches uranium at a specified capacity. \\
Facility & Fab$^*$ & This facility fabricates fuel material based on separated streams. \\
Facility & Reactor$^*$ & A reactor model that handles batch refueling, based on pre-determined recipes of compositions. It requests any number of input fuel types and transmutes them to static compositions upon discharge.\\
Facility & Separations$^*$ & This facility separates an incoming material into specified streams. \\
Facility & Sink & This facility is capable of accepting a finite or infinite quantity of some commodity produced in the simulation. \\
Facility & Source & This facility generates material of the composition and commodity type specified as input.  \\
Institution & ManagerInst & The manager institution manages production of commodities among its facilities by building new ones as needed. \\
Institution & DeployInst &  This institution deploys specific facilities as defined explicitly in the input file. \\
Region & GrowthRegion & This region determines whether there is a need to meet a certain capacity (as defined via input) at each time step. \\
\hline
\end{tabularx}
\caption{The Archetypes in \Cycamore seek to cover a large range of simple
simulation use cases \cite{carlsen_cycamore_2014}. Facilities added in version
1.3 are marked with $^*$.}
\label{tab:cycamore}
\end{table}

@@

\subsection{Quality Assurance}
% Organize this section according to major topics
% give each topic a section heading in boldface.
% try to cover the major common points :
%
% problem design
% methods of measurement
% supporting models
% supporting data
% simulations run
% results

% Just write the section headings for each part and indicate what goes in that
% section with words :
%
% heading
% figures (with captions)
% schematics (with captions and footnotes)
% equations
% tables

% What does it mean?
% What did I actually test?
% What were the results?
% Did the work yield a new method?
% Did the work yield new knowledge?
% What measurements did I make?
% How were these measurements characterized?
% What methods were used?
% What were the results?
% How were the measurements made and characterized?

Simulation science - like experimental science - must distinguish trustworthy
results from untrustworthy ones.
Charles Babbage famously articulated this, ``On two occasions I have been asked,
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right
answers come out?' ... I am not able rightly to apprehend the kind of confusion
of ideas that could provoke such a question.'' \cite{babbage_passages_2011}.
A simulator must assure correctness and guard against the \emph{garbage
in, garbage out} phenomenon.

Multiple strategies, collectively known as \emph{\gls{QA}}, have
been invented to mitigate structural and algorithmic errors
in software. These include \emph{\gls{VV}}
\cite{boehm_software_1989}, testing, and others.

@@

Verification may be defined as the question, ``Is \Cyclus being built correctly?''
To answer this question, \Cyclus relies on software development best practices
such as testing,
documentation, version control, style guidelines, and continuous integration to
ensure reliability and reproducibility. These verification techniques are
discussed individually in the sections that follow.

Validation, in contrast,  may be defined as the question,
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal
validation is not possible. Nonetheless, code-to-code comparisons with fuel cycle
simulators that use other modeling paradigms are underway
\cite{huff_extensions_2014}.

Sections \ref{sec:qa-testing}-\ref{sec:qa-ci} discuss in greater detail the software
development components that comprise the \Cyclus verification strategy, which 
allows the simulation kernel and physics models to be tested explicitly and 
separately.
Each of these approaches on its own is a valuable addition to \gls{QA} but it cannot be the
entire answer to the requirements imposed by \gls{NQA1}. Taken together and strictly
adhered to, they present a fortress to protect against poorly designed or otherwise undesirable code.

@@

Integration tests combine multiple elements of the
\Cyclus interface and test that they work correctly with each other.  By analogy,
simply because the gears (units) are made correctly does not imply that the
clock (integration) will run smoothly, run at all, or give the correct time.
In \Cyclus and \Cycamore, integration tests are performed by running sample
simulations for scenarios that are sufficiently simple to predict the results
and verifying that results match those predictions. A set of standard input
files are run, then the output is inspected and compared via Nose
\cite{pellerin_nose_2007}, a Python test framework.  In this way, \Cyclus code
units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that
major \Cyclus components are functioning as expected.

@@

\subsubsection{Documentation}

@@

\subsubsection{Version Control}

Version control preserves the development history, or \emph{provenance}, of
source code files. \Cyclus uses a well-established \emph{distributed version
control} tool called git \cite{software_freedom_conservancy_git_2014} for this
purpose.  Distributed version control allows every
user to have complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
One feature that git performs exceptionally well is \emph{branching}.
Branches are distinct pathways in the history that diverge from a mainline source
tree and then may be \emph{merged} back in. Multiple simultaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an
accompanying message. Thus,
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.

\Cyclus uses a strategy known as \emph{git flow}
\cite{kalliamvakou_code-centric_2014} to manage topical branches where
individuals develop code in parallel. All proposed software changes from
topical branches must have sufficient successful tests and comprehensive
documentation to be allowed into \Cyclus.  As an example, a schematic of the
development stages between reporting a bug and merging the fix is shown in
Figure \ref{fig:gitprocess}.

\begin{figure}[htbp]
\begin{center}
\includegraphics[height=0.6\textheight]{./images/gitprocess}
\end{center}
\caption{A version-control enabled process for finding, reporting, and fixing a bug
can take place within an issue tracker alongside multiple stages of software
testing and review. In the \Cyclus case, the final review stage includes unit,
regression, and integration testing as well as manual code inspection.}
\label{fig:gitprocess}
\end{figure}

The act of proposing a change is known as a \emph{pull request}. The main \Cyclus repository is
hosted remotely on the GitHub website \cite{dabbish_social_2012}. The online
mechanism for pull requests allows for code review by a member of the \Cyclus
core team, separate from the author, prior to any change being included. Non-members
of the \Cyclus development team are allowed and encouraged to submit and review
pull requests. However, only members of the \Cyclus core team are allowed to
merge in any changes and only after the \gls{QA} standards have been met. This
step has been repeatedly shown to improve code quality \cite{cohen_modern_2010}.

\subsubsection{Style Guide}

@@

\subsubsection{Continuous Integration}
\label{sec:qa-ci}

\emph{Continuous integration (CI)} is the idea that software should be tested and validated
as it is being developed, rather than as a final stage in a longer development
cycle.  Under \gls{CI}, every pull request (proposed code change) is tested independently immediately after
it is proposed. Such tests are run on all officially supported platforms.
\Cyclus uses a \gls{CI} platform called Polyphemus
\cite{scopatz_polyphemus_2014}. Polyphemus serves as an intermediary between GitHub pull requests on the front-end
and temporary \Cyclus servers on the back-end. These servers are hosted by
the Build \& Test Laboratory (BaTLab) \cite{uw_batlab_team_batlab_2014} at the University of
Wisconsin-Madison. Whenever a pull request is created, Polyphemus performs
the following actions:

\begin{enumerate}
    \item Copies \Cyclus with the proposed new code to BaTLab,
    \item Initializes Linux and Mac OSX servers,
    \item Builds \Cyclus on all platforms,
    \item Runs the complete \Cyclus test suite,
    \item Reports whether or not the above steps succeeded.
\end{enumerate}

Since these steps are performed for all incoming code, it is easy for the
\Cyclus core team to determine whether or not an individual pull request
actually works. This helps identify build and test problems prior to
broken code being allowed into the develop branch. \gls{CI} is a necessary
but not sufficient addition to the \Cyclus \gls{QA}
system: it keeps bad code out of \Cyclus. However, \Cyclus will always
require human eyes and hands to let good code in.

>>>>

\begin{figure}[htbp!]
\begin{center}
\includegraphics[width=0.5\textwidth]{./images/modifiedopen.eps}
\end{center}
\caption{The \Cyclus framework enables fully open, partially open, and fully
closed collaborations\cite{carlsen_cyclus_2014}.}
\label{fig:modifiedopen}
\end{figure}

@@

\begin{figure}[htbp!]
\begin{center}
\includegraphics[width=\textwidth]{./images/compositions}
\end{center}
\caption{
A simple decay line cache holding compositions A, B, C, and a
yet-uncomputed composition D.  B comes from decaying A 1 time step.  C comes
from decaying B 2 time steps, etc.  Black represents the cache for this
particular composition chain. Blue indicates decay operations that can be
satisfied by cache lookups.   If A needs to be decayed 1 time step (A +1), a
quick lookup returns the previously computed B.  Decaying B 3 time steps
requires a decay calculation to compute a new composition D, but subsequent
requests such as decaying A 4 time steps will not require any recalculation.
}
\label{fig:compositions}
\end{figure}

@@

\begin{table}[H]
\centering
\begin{tabularx}{\textwidth}{|r|l|X|}
\hline
\textbf{Entity} & \textbf{Archetype} & \textbf{Functionality} \\
\hline
Facility & EnrichmentFacility & This facility enriches uranium at a specified capacity. \\
Facility & Fab$^*$ & This facility fabricates fuel material based on separated streams. \\
Facility & Reactor$^*$ & A reactor model that handles batch refueling, based on pre-determined recipes of compositions. It requests any number of input fuel types and transmutes them to static compositions upon discharge.\\
Facility & Separations$^*$ & This facility separates an incoming material into specified streams. \\
Facility & Sink & This facility is capable of accepting a finite or infinite quantity of some commodity produced in the simulation. \\
Facility & Source & This facility generates material of the composition and commodity type specified as input.  \\
Institution & ManagerInst & The manager institution manages production of commodities among its facilities by building new ones as needed. \\
Institution & DeployInst &  This institution deploys specific facilities as defined explicitly in the input file. \\
Region & GrowthRegion & This region determines whether there is a need to meet a certain capacity (as defined via input) at each time step. \\
\hline
\end{tabularx}
\caption{The Archetypes in \Cycamore seek to cover a large range of simple
simulation use cases \cite{carlsen_cycamore_2014}. Facilities added in version
1.3 are marked with $^*$.}
\label{tab:cycamore}
\end{table}

@@

\subsection{Quality Assurance}

% topics:
%  - version control
%  - PR policy
%  - style guide
%  - testing
%  - CI

To garner the trust of a broad user and archetype developer community, the
\Cyclus project must implement a strategy to assure the ongoing quality of the
software it provides.  Multiple strategies, collectively known as
\emph{\gls{QA}}, have been developed by the scientific software development
community to mitigate structural and algorithmic errors in software. These
include \emph{\gls{VV}} \cite{boehm_software_1989}, testing, and others.

@@

Validation of simulators like \Cyclus, that are intended to give forecasts of
system behavior in uncertain futures, is not well defined as there are no
experimental benchmarks upon which to rely.  Instead, code-to-code comparisons
with fuel cycle simulators that use other modeling paradigms are underway as
the best approach to establish confidence that \Cyclus produces correct
answers. \cite{huff_extensions_2014}

Verficiation of \Cyclus implementation relies on software development best
practices such as version control, testing, continuous integration,
documentation, and style guidelines to ensure reliability and
reproducibility. Sections \ref{sec:qa-vc}-\ref{sec:qa-review} discuss in
greater detail the software development components that comprise the \Cyclus
verification strategy, which allows the simulation kernel and physics models
to be tested explicitly and separately. Each of these approaches on its own is
a valuable addition to \gls{QA} but it cannot be the entire answer to the
requirements imposed by \gls{NQA1}. Taken together and strictly adhered to,
they present a fortress to protect against poorly designed or otherwise
undesirable code.

\subsubsection{Version Control}
\label{sec:qa-vc}

Automated version control, provided by the well-established \emph{distributed
  version control} tool git \cite{software_freedom_conservancy_git_2014}, is
at the heart of the QA strategy because it records the full history of any
change, along with metadata such as the author, a timestamp, and an
accompanying message. This makes it possible to identify when changes were
introduced, how they related to other changes, and who made those changes.  If
necessary, it also facilitates reversing individual changes from a long and
intricate set of changes.  Version control also enables a code review process
descried in section \ref{sec:qa-review} below.

@@

Integration tests combine multiple elements of the
\Cyclus interface and test that they work correctly with each other. 
In \Cyclus and \Cycamore, integration tests are performed by running sample
simulations for scenarios that are sufficiently simple to predict the results
and verifying that results match those predictions. A set of standard input
files are run, then the output is inspected and compared via Nose
\cite{pellerin_nose_2007}, a Python test framework.  In this way, \Cyclus code
units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that
major \Cyclus components are functioning as expected.

@@

\paragraph{Continuous Integration}

\emph{Continuous integration (CI)} is the idea that software should be tested
and validated as it is being developed, rather than as a final stage in a
longer development cycle.  Under \gls{CI}, every pull request is automatically
tested independently immediately as it is proposed. The results of this
testing are shown to code reviewers in the GitHub interface.  The \Cyclus
project currently uses the CircleCI service for continuous integration.
Whenever a pull request is created, CircleCI builds a version of the \Cyclus
source code that includes the requested changes and runs the complete test
suite, reporting back whether or not those steps were successful. If \gls{CI}
was not successful, the code author(s) must first identify and resolve the
problems.

Since these steps are performed for all incoming code, it is easy for the
\Cyclus core team to determine whether or not an individual pull request
actually works. This helps identify build and test problems prior to broken
code being allowed into the main software development branch.

\subsubsection{Code Review}
\label{sec:qa-review}

Autmoated testing including \gls{CI} is a necessary but not sufficient
component of the \Cyclus \gls{QA} system: it keeps bad code out of
\Cyclus. However, \Cyclus will always require human eyes and hands to let good
code in.

The main \Cyclus repository is hosted remotely and publicly on the GitHub
website \cite{dabbish_social_2012}, in part because it provides tools that
facilitate a culture of frequent and thorough code review.  A number of
policies exist to ensure that a proposed set of changes, known as a \emph{pull
  request}, adhere to the projects \gls{QA} standards. Every pull request must
be reviewed and accepted by a member of the \Cyclus core team that was not
involved in authoring the changes.  In addition to reviewing the algoirthmic
design in the source code, the reviewer relies on tests to ensure correct
behavior, and requires the authors to adhere to the style guide and provide
sufficient documentation.  Only after the \gls{QA} standards have been met are
the proposed changes merged into the software. This step has been repeatedly
shown to improve code quality \cite{cohen_modern_2010}.

Once \gls{CI} is successful, a code reviewer inspects not only the
changes that are proposed to the software itself, but also the changes that
have been proposed to tests.  It is important that the subset of tests that
are relevant to the software change has a scope and completeness that will
ensure that incorrect behavior will be automatically detected.  If there are
sufficent tests and \gls{CI} is successful, the reviewer can proceed to
considering the actual software changes themselves.

Because of their role in facilitating code review during the change process,
as well as their long term benefit to the maintainability of the project,
documentation and coding style are also reviewed as part of this process.

\paragraph{Documentation}

@@

\paragraph{Style Guide}

@@
